Canonical rings of $\bq$-divisors on minimal rational surfaces

### 2016-08-04 surfaces-write-up/surface-write-up.tex
modified hilbert rational surfaces

We expect the work in this paper will help calculate a presentation of certain rings of
Hilbert 
modular forms and Siegel modular forms. Recall that Hilbert and Siegel 
modular forms are two generalizations of modular forms to higher dimensions,
as described in \cite{geer:siegel-modular} and \cite{bruinier:hilbert-modular}.
Voight and Zureick-Brown show that 
one can translate the analytic category of 1-dimensional complex orbifolds to
the algebraic category of stacky curves \cite[Proposition 6.1.5]{vzb:stacky}. 
We expect that one can generalize this result to yield an equivalence of
categories between an appropriate subcategory of $m$-dimensional orbifolds and
a subcategory of $m$-dimensional algebraic spaces.
In particular, this would allow us to translate rings of Hilbert modular forms
and Siegel modular forms whose coarse space is a projective space or a
Hirzebruch surface to canonical rings on the corresponding algebraic surfaces.
Thereby, if our results could be extended to all rational surfaces,
they would give a bound on generators and relations for such
rings of Hilbert and Siegel modular forms parameterized by rational
Hilbert and Siegel modular surfaces.
This would be interesting because such modular surfaces tend to be
immensely complicated: ``Already for $g = 2$, the boundary of this fundamental domain is complicated; Gottschling found that it possesses $28$ boundary pieces'' \cite[p.\ 7]{geer:siegel-modular}.
Since these varieties are so complicated, it is difficult
to give explicit examples in which we can completely work out the canonical
ring of a modular surface.
Although we did not obtain a bound on the degree of generators and relations
for general rational surfaces, the restricted class of rational varieties we consider still required significant work.

>>>>

\subsection{Modular Forms}
The work in this paper was motivated by 
potential applications to
calculating a presentation of certain rings of
Hilbert 
modular forms and Siegel modular forms. 
Recall that Hilbert and Siegel 
modular forms are two generalizations of modular forms to higher dimensions,
as described in \cite{geer:siegel-modular} and \cite{bruinier:hilbert-modular}.
Voight and Zureick-Brown show that 
one can translate the analytic category of 1-dimensional complex orbifolds to
the algebraic category of stacky curves \cite[Proposition 6.1.5]{vzb:stacky}. 
If one could generalize this result to yield an equivalence of
categories between an appropriate subcategory of $m$-dimensional orbifolds and
a subcategory of $m$-dimensional algebraic spaces,
this would allow us to translate rings of Hilbert modular forms
and Siegel modular forms whose coarse space is a projective space or a
Hirzebruch surface to canonical rings on the corresponding algebraic surfaces.
Thereby, if our results could be extended to all rational surfaces,
they would give a bound on generators and relations for such
rings of Hilbert and Siegel modular forms parameterized by rational
Hilbert and Siegel modular surfaces.
This would be interesting because such modular surfaces tend to be
immensely complicated: ``Already for $g = 2$, the boundary of this fundamental domain is complicated; Gottschling found that it possesses $28$ boundary pieces'' \cite[p.\ 7]{geer:siegel-modular}.
Since our results only apply to minimal rational surfaces, and because the rings of Hilbert and Siegel modular forms are so complex, we were unable
to use our work to compute the canonical ring of an explicit modular surface.
Although we did not obtain a bound on the degree of generators and relations
for general rational surfaces, the restricted class of rational varieties we consider still required significant work.
